# Setting reference levels {#sec-ref}

Up to this point I used reference levels (RL) drawn from my
own personal and subjective intuition. There might be more objective ways to 
do this, or at least more precise and defendable ways. What we have decided on 
is to create a se of calibrartion maps, with varying amounts of trenching,
and then several experts will look at these and judge individually how severe
they believe a given density of trenches will be for the mire ecosystem. 
We need (minimum) three RLs. 

X0 is the variable value (i.e. the frequence of 
10x10 grid cells with trenching inside each 100x100 cell) that corresponds to a
completely destroyed ecosystem. 
A further deterioration of the variable value should now result in any further
deterioratio of ecosystem condition.

X100 is the RL at the reference condition (RC; _pristine condition_). 
This RL is quite intuitive and is set to zero.

X60 represents the variable value at the boundary between good and deteriorated 
ecosystem condition. We have decided to anchor this RL theoretically to the
point where trenching is so severe that the mire ecosystem will inevitably turn 
into terrestrial land, meaning it will not be wtland anymore (but it can be peatland).

```{r setup}
#| include: false
#| warning: false

library(tidyverse)
library(stars)
library(ggmap)

# Get the geometry for the 100x100 grid cells
# with a column for the frequency of 10x10m grid cells that have trenches
grid100_freq4 <- readRDS("data/grid100_freq4.rds")

# Get the trenches
load("data/study-area-files.RData")
```

```{r sateliteImages}
#| eval: false
#| code-fold: true
#| code-summary: "Set up Google API, download and save satelite images"


# I used this to save the API key to my syst env
# ggmap::register_google(key = "", write = TRUE)

# get center coordinate
e <- st_as_sfc(st_bbox(grid100_freq4)) |>
  st_transform(4326) |>
  st_centroid() |>
  st_coordinates() |>
  as.numeric()
names(e) <- c("lon", "lat")
  
satImg <- ggmap:: get_googlemap(
  center = e, 
  zoom = 14,
  maptype = "satellite"
  )

saveRDS(satImg, "data/satImg.rds")

```

```{r readSatImg}

satImg <- readRDS("data/satImg.rds")
#plot(satImg)

one <- grid100_freq4 |>
  slice_max(freq) |>
  st_transform(3857)


# Hackish function to fix bbox in the ggmap object
# https://stackoverflow.com/questions/47749078/how-to-put-a-geom-sf-produced-map-on-top-of-a-ggmap-produced-raster
ggmap_bbox <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))

  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))

  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}

# Use the function:
map <- ggmap_bbox(satImg)

ggmap(ggmap = map)+
  coord_sf(crs = st_crs(3857))+
  geom_sf(data = one, inherit.aes = F)
```

