# Study area and trenching data {#sec-area}

In this project we will use data from Levanger in central Norway. This area was part of the pilot study to identify mire trenches in Norway based on LiDAR (see @sec-background).

```{r setup}
#| warning: false
#| error: false
#| include: false
library(sf)
library(tidyverse)
library(basemaps)
options(scipen = 999)
```

```{r data}
#| code-fold: true
#| code-summary: "Read data" 

path <- "P:/41201785_okologisk_tilstand_2022_2023/data/grofter/Kolstad"
lev <- sf::read_sf(paste0(path, "/LevangerGroftPolygon.shp")) |>
  sf::st_make_valid()

```

```{r stats}
#| code-fold: true
#| code-summary: "Calculate study extent"

lev_bb <- sf::st_bbox(lev)
myx <- unname(round(lev_bb["xmax"] - lev_bb["xmin"]))
myy <- unname(round(lev_bb["ymax"] - lev_bb["ymin"]))
```

The study area is `r myx` ð–· `r myy` meters, or `r round(myx*myy/10^6)` km^2^.

```{r basemapSetup}
#| code-fold: true
#| code-summary: "Set up basemap"

basemaps::set_defaults(map_service = "osm", map_type = "topographic")
lev2 <- lev |>
  st_transform(3857)
```

```{r fig-map1}
#| code-fold: true
#| code-summary: "Make map"
#| fig-cap: |
#|   "Map of study area in Levanger, central Norway, showing trenched mires i purple. 
#|   Note that the colored polygon borders exagerrate the area of the trenches."
#| cache: true
#| message: false

ggplot() + 
  basemap_gglayer(sf::st_bbox(lev2)) +
  scale_fill_identity() + 
  coord_sf() +
  geom_sf(data = lev2,
          fill = "purple",
          color = "purple") +
  theme_bw()
```

```{r zoombbox}
#| code-fold: true
#| code-summary: "Creat new bbox"
#| warning: false

bbox_new <- st_bbox(lev2) # current bounding box

xrange <- bbox_new$xmax - bbox_new$xmin # range of x values
yrange <- bbox_new$ymax - bbox_new$ymin # range of y values

bbox_new[1] <- bbox_new[1] + (0.3 * xrange) # xmin - left
bbox_new[3] <- bbox_new[3] - (0.5 * xrange) # xmax - right
bbox_new[2] <- bbox_new[2] + (0.4 * yrange) # ymin - bottom
bbox_new[4] <- bbox_new[4] - (0.5 * yrange) # ymax - top

lev2_crop <- lev2 |>
  sf::st_crop(bbox_new)

#bbox_new <- bbox_new[c(1,3,2,4)]
```

```{r fig-map2}
#| code-fold: true
#| code-summary: "Create map"
#| cache: true
#| fig-cap: |
#|   "Zoomed in version of the previsous map showing systematic trenching,
#|   but also more sporadic trenches that might be errors in the model."

ggplot() + 
  basemap_gglayer(bbox_new) +
  scale_fill_identity() + 
  geom_sf(data = lev2_crop,
          fill = "purple",
          color = "purple") +
  theme_bw()+
  coord_sf()
```

The trenches model originally results in a raster with a continuous probability value for the occurence of a trench in each 1 ð–· 1 m cell. What I show above (@fig-map1, @fig-map2) is a version of this data where a threshold value for the probabilities turn it into presence-absence data in the shape of polygons.

## Mire data
